### 2021-04-04 Chapter-4/Chapter-4.tex
Updates from Overleaf

%% to deal with side channels, we can protect or detect them and detection is better
%%arious countermeasures have been proposed to defend against software-based
%%side-channel attacks. Hardware-level solutions, such as reducing shared
%%resources, adopting oblivious RAM, and using transnational
%%memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need new
%%hardware features and changes in modern complex computer systems, which are
%%impractical and hard to adopt in reality. Therefore, a promising and
%%universal direction is software countermeasures, detecting and eliminating
%%side-channel vulnerabilities from code base.

Many side-channel attacks originate
in two code patterns: data flow from secrets to load
addresses and data flow from secrets to branch conditions. We refer to them as
secret-dependent data accesses and control flows, respectively.

@@

We assume that an attacker shares the same hardware platform with the target.
The attacker attempts to retrieve sensitive information through address-based
side-channel attacks. The attacker has no direct access to the target's memory or cache,
but it can probe its memory or cache at each program point. In reality, the
attacker will face many possible obstacles such as the noisy observations 
of the memory or cache. However, for this project, we assume
the attacker has noise-free observations as in previous work~\cite{203878,182946,Brotzman19Casym}. 
The threat model captures most cache-based and address-based side-channel attacks. 
We only consider deterministic programs and assume an attacker 
has access to the source code or binary executable of the target program.

@@

\begin{figure}[h!]
\centering
\begin{lstlisting}[xleftmargin=.03\textwidth,xrightmargin=.01\textwidth]
uint8_t key[2], t1, t2;
get_key(key);              // 0 <= key[0], key[1] < 256
t1 = key[0] + key[1];
t2 = key[0] - key[1];
if (t1 < 4){               // leakage site A
    foo();    
}                          
if (t2 > 0){               // leakage site B     
    doo();    
}                          
\end{lstlisting}
\vspace*{-1pt}
    \caption{Side-channel leakage}
    \label{fig:side-channel}
\end{figure}
In this paper we assume an attacker can observe the secret-dependent control-flows in Figure~\ref{fig:side-channel}.
Therefore, an attacker can have two different observations for each leak site
depending on the value of the $\mathit{key}$: $A$ for function \textsf{foo} is executed, 
$\neg A$ for function \textsf{foo} is not executed, $B$ for function \textsf{doo} is
executed, and $\neg B$ for function \textsf{doo} is not executed. Now the
question is how much information can be leaked from the above code if an
attacker knows which branch is executed?

\begin{table}[ht]
    \centering\small\footnotesize
    \caption{The distribution of observation}\label{shtable}
    \vspace*{-0pt}
%    \resizebox{\columnwidth}{!}{
    \begin{tabular}{l|cc|cc}
        \hline

@@

 \begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.03\textwidth,xrightmargin=.01\textwidth]
char key[9] = input();
if(strcmp(key, "password"))   // leakage site C
    pass();                   // branch 1
else
    fail();                   // branch 2
\end{lstlisting}
\vspace*{-10pt}
    \caption{A dummy password checker}
    \label{fig:password-checker}
\end{figure}

@@

However, prior static approaches cannot precisely capture the amount of leakage. According to the definition of Shannon entropy, the leakage will be
$\frac{1}{2^{64}}*\log_{2}\frac{1}{2^{64}} + \frac{2^{64}-1}{2^{64}}
*\log_{2}\frac{2^{64}-1}{2^{64}} \approx 0$ bits. Max-entropy is defined from the
number of possible observations. Because the program has two
branches, tools based on max-entropy will report the code has a $\log_2{2} = 1$
bit leakage.
Both approaches fail to tell how much information is leaked during the execution
precisely. The problem with existing methods is that they are static-based so
input values and real runtime information are neglected by their approaches. 
They assume an attacker runs
the program multiple times with many different or random sensitive inputs. As
shown in Figure~\ref{fig:gap}(a), previous models, both Shannon entropy and max
entropy, give an ``average'' estimate of the information leakage. However, it is
not the typical scenario for an adversary to launch a side-channel attack. When
a side-channel attack happens, the adversary wants to retrieve the sensitive
information, in which case the sensitive information is fixed (e.g., AES keys).
The adversary will run the attack over and over again with fixed input and 
guess the value bit by
bit (e.g., Kocher's timing attacks~\cite{kocher1996timing}), as in Figure~\ref{fig:gap}(b). We want to have a
theory for dynamic analysis that if the theory says an attack leaks $x$ bits of
secret information from a side-channel vulnerability, then $x$ should be useful
in estimating the sensitive level of the vulnerability. However, the above
methods all fail in real attack models. This is the first challenge we face
\textbf{(Challenge C1)}.

\begin{figure}
    \centering
    \includegraphics[width=.65\columnwidth]{./figures/chapter4/RA.pdf}
    \caption{The gap between the real attack and previous models}\label{fig:gap}
\end{figure}

@@

>>>>

@@

The threat model is similar to our previous work. That is, we assume 
an attacker can share some hardware resource with the victim. The attacker has
the ability to probe the victim at any time of the execution. We also assume 
an attacker has noise-free observations. The attacker has the access to the victim's
binary executable. The threat mode is similar to 
the previous work~\cite{203878,182946,Brotzman19Casym} and can catch most of 
the address-based side-channel attacks in the literature.

@@

\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
uint8_t key[2], t1, t2;
get_key(key);              // 0 <= key[0], key[1] < 256
t1 = key[0] + key[1];
t2 = key[0] - key[1];
if (t1 < 4){               // leakage site A
    foo();    
}                          
if (t2 > 0){               // leakage site B     
    doo();    
}                          
\end{lstlisting}
    \vspace*{-1pt}
    \caption{Side-channel leakage}
    \label{fig:side-channel}
\end{figure}
In this paper we assume an attacker can observe the secret-dependent control-flows in Figure~\ref{fig:side-channel}.
Therefore, an attacker can have two different observations for each leak site
depending on the value of the $\mathit{key}$: $A$ for function \textsf{foo} is executed,
$\neg A$ for function \textsf{foo} is not executed, $B$ for function \textsf{doo} is
executed, and $\neg B$ for function \textsf{doo} is not executed. Now the
question is how much information can be leaked from the above code if an
attacker knows which branch is executed?

\begin{table}[ht]
    \centering\small\footnotesize
    \caption{The distribution of observation}\label{shtable}
    %    \resizebox{\columnwidth}{!}{
    \begin{tabular}{l|cc|cc}
        \hline

@@

\begin{figure}[h!]
    \centering
    \begin{lstlisting}[xleftmargin=.2\textwidth, xrightmargin=.2\textwidth]
char key[9] = input();
if(strcmp(key, "password"))   // leakage site C
    pass();                   // branch 1
else
    fail();                   // branch 2
\end{lstlisting}
    \vspace*{-10pt}
    \caption{A dummy password checker}
    \label{fig:password-checker}
\end{figure}

@@

However, prior static approaches cannot precisely capture the amount of leakage. According to the definition of Shannon entropy, the leakage will be
$\frac{1}{2^{64}}*\log_{2}\frac{1}{2^{64}} + \frac{2^{64}-1}{2^{64}}
    *\log_{2}\frac{2^{64}-1}{2^{64}} \approx 0$ bits. Max-entropy is defined from the
number of possible observations. Because the program has two
branches, tools based on max-entropy will report the code has a $\log_2{2} = 1$
bit leakage.
Both approaches fail to tell how much information is leaked during the execution
precisely. The problem with existing methods is that they are static-based so
input values and real runtime information are neglected by their approaches.
They assume an attacker runs
the program multiple times with many different or random sensitive inputs. As
shown in Figure~\ref{fig:gap}(a), previous models, both Shannon entropy and max
entropy, give an ``average'' estimate of the information leakage. However, it is
not the typical scenario for an adversary to launch a side-channel attack. When
a side-channel attack happens, the adversary wants to retrieve the sensitive
information, in which case the sensitive information is fixed (e.g., AES keys).
The adversary will run the attack over and over again with fixed input and
guess the value bit by
bit (e.g., Kocher's timing attacks~\cite{kocher1996timing}), as in Figure~\ref{fig:gap}(b). We want to have a
theory for dynamic analysis that if the theory says an attack leaks $x$ bits of
secret information from a side-channel vulnerability, then $x$ should be useful
in estimating the sensitive level of the vulnerability. However, the above
methods all fail in real attack models. 
\begin{figure}
    \centering
    \includegraphics[width=.65\columnwidth]{./figures/chapter4/RA.pdf}
    \caption{The gap between the real attack and previous models}\label{fig:gap}
\end{figure}

@@

\subsubsection{Math Notations}
We assume a program ($\beta$) has $K$ as its sensitive input. $K$ should be 
a finite set of keys. The program also takes known messages $M$ as its input. 
During an AES encryption, for example, 
$\beta$ is the encryption function. $K$ is the set of all possible AES keys, 
and $M$ represents the set consisting of all plaintext messages to be encrypted. In a real execution, an adversary may have 
some observations ($O$) of the program. This paper only
uses secret-dependent control flows and secret-dependent data 
accesses as observations.

With the above definition, we have the following mapping between $\beta$,
$K$, $M$, and $O$:
%\vspace*{-0pt}
\begin{displaymath}
    \beta(K, M) \rightarrow O
\end{displaymath}

We model a side-channel in the following way. An adversary does not have
access to $K$, but he knows $\beta$, $M$, and $O$. For one execution of a
deterministic program, once $k \in K$ and $m \in M$ are fixed, the observation
($o \in O$) should also be determined. An attacker knows $\beta$, $o$,
and $m$. The attacker wants to infer the value of $k$. We use $K^o$ to denote
the set of possible $k$ values that produce the same observation: $K^o = \{ k \in K \, |\, \beta(k, m) \rightarrow o\}$

Then the problem of quantifying the amount of leaked information can be
restated as the following question:

\emph{How much uncertainty of $K$ is reduced if an attacker knows $\beta$, $m$, and $o$?}
